---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# survey.workflow

<!-- badges: start -->
[![R-CMD-check](https://github.com/DSCI-310-2025/pkg.drugage/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/DSCI-310-2025/pkg.drugage/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/DSCI-310-2025/pkg.drugage/graph/badge.svg)](https://app.codecov.io/gh/DSCI-310-2025/pkg.drugage)
<!-- badges: end -->

The goal of survey.workflow is to ...

## Installation

You can install the development version of survey.workflow from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("DSCI-310-2025/survey.workflow")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(survey.workflow)
## basic example code
```

## License

This project is dual-licensed under the following licenses:

- **[Creative Commons Zero v1.0 Universal (CC0 1.0)](https://creativecommons.org/publicdomain/zero/1.0/)**
  You can copy, modify, distribute and perform the work, even for commercial purposes, all without asking permission.

- **[MIT License](LICENSE.md)**  
  A short and permissive software license. You can use it in proprietary software provided that all copies include the license terms and the copyright notice.
